PhD Title \\ PhD Thesis

### 2024-07-24 Latex_files/Thesis/chapters/chapter 5/main.tex
updates

\chapter{Detection and Characterisation of rotational spherical aggregate rotational dynamics}
\label{chapter:simulated_detection}
As outlined in the end of Chapter 4, one of the difficulties in characterising 
interactions with asymmetric objects is the coupled motion between translation 
and rotation. This can be somewhat ignored if the rotational motion is periodic,
like when the power spectra of elliptical polystyrene particles was fitted by 
Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, however in the case where 
rotational motion is stochastic the problem is more complex. For example, when a
novel trap characterisation technique was implemented by Saffron \textit{et al}
\cite{BarZiv1997, Meller1998}, they were able to use dynamic light scattering to
characterise both the axial and lateral trap stiffness acting on microspheres. The
only drawback admitted to in their work was that the technique was constrained to
isotropic scatters as their theoretical model for describing the auto-correlation
function was predicated on the fact that any variations in the signal are due to
the particles translational motion within the confines of a cylindrical trap.

@@

\subsection{Coordinate System}
In the case of our probe beam the +z direction points from the surface of the probe directly to the centre of diffusion in our dimer, with the +y direction pointing towards what would be our trapping beam, and the +x direction pointing into the page. The origin of our coordinate system is fixed on the centre of the focus of our trapping beam, meaning as our dimer's centre of diffusion moves the origin is kept constant.
\subsection{Dimer}
The dimer is defined by two spheres, in the trapping frame this would be orientated by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as the large sphere is furthest from the trapping beam. We scale the position of each sphere by the factor L:
\begin{align}
	L = \frac{1}{k} = \frac{\lambda}{2\pi}
\end{align}

Where $\lambda$ is the wavelength of light and is given in nanometres, this means every position must be also provided in nanometres.
\subsection{Beam}
Our probe beam is rather simple to define, we firstly say that the probe is a plane wave by setting the Gaussian beam parameter $C_B$ to 0. We also say the beam is un-polarized by defining the Stokes vector as $[1,0,0,0]$.

\subsection{Detectors and Pixels}

Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted later for testing), with the position of the detector being defined by the polar and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
\begin{align}
	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
\end{align}
So if you were to place the fibers in the trapping frames x-y plane, in the probe beam frame will only have x and z components. The orientation of the detector is just the -ve of its position and not scaled by the distance term r, the orientation being the vector from the origin of the detector to the center of the dimer.

Pixels can be thought of as points lying on the surface of the detector, by getting the spherical coordinates of each pixel we can compute the intensity at each point on the detector, allowing us to determine the average intensity on a given detector. The pixels are distributed using a sunflower distribution given by:

@@

This allows us to define the surface of any detector perpendicular to the origin of our coordinate system, we can freely choose the size and resolution of our fibre's. For every pixel we define the spherical coordinates relative to our origin, this is added to a list of $\theta, \phi$ pairs that are evaluated by MSTM.

\section{Interpretation of scattering data into orientation estimates}
MSTM is capable of evaluating the scattering characteristics of a given target either in a fixed orientation or using an averaging of the particles orientation. While the latter method is satisfactory for simulation purposes it

>>>>

\chapter{Detection and Characterisation of rotational spherical aggregate rotational dynamics}
\label{chapter:simulated_detection}
As outlined in the end of Chapter 4, one of the difficulties in characterising 
interactions with asymmetric objects is the coupled motion between translation 
and rotation. This can be somewhat ignored if the rotational motion is periodic,
like when the power spectra of elliptical polystyrene particles was fitted by 
Yogesh \textit{et al} \cite{Yogesha2011PreciseCO}, however in the case where 
rotational motion is stochastic the problem is more complex. For example, when a
novel trap characterisation technique was implemented by Saffron \textit{et al}
\cite{BarZiv1997, Meller1998}, they were able to use dynamic light scattering to
characterise both the axial and lateral trap stiffness acting on microspheres. The
only drawback admitted to in their work was that the technique was constrained to
isotropic scatters as their theoretical model for describing the auto-correlation
function was predicated on the fact that any variations in the signal are due to
the particles translational motion within the confines of a cylindrical trap 
\cite{BarZiv1997}. However, it is somewhat naive to assume that the orientational 
change

@@

\subsection{Coordinate System}
In the case of our probe beam the +z direction points from the surface of the 
probe directly to the centre of diffusion in our dimer, with the +y direction 
pointing towards what would be our trapping beam, and the +x direction pointing 
into the page. The origin of our coordinate system is fixed on the centre of 
the focus of our trapping beam, meaning as our dimer's centre of diffusion moves 
the origin is kept constant.

\subsection{Beam}
The probe is a defined as a plane wave by setting the Gaussian beam parameter 
$C_B$ to 0. We also say the beam is un-polarized by defining the Stokes vector 
as $[1,0,0,0]$. The plane wave in \textit{mstm} is given by \eqref{eq:incident}.

\subsection{Dimer}
The dimer is defined by two spheres, in the trapping frame this would be orientated 
by $s=[0,0,1]$ however in our scattering set up this is rotated to $s=[-1,0,0]$ as 
the large sphere is furthest from the trapping beam. We scale the position of each 
sphere by the factor L:
\begin{align}
	L = \frac{1}{k} = \frac{\lambda}{2\pi}
\end{align}

Where $\lambda$ is the wavelength of light and is given in nanometres, this means 
every position must be also provided in nanometres.

\subsection{Detectors and Pixels}
Each detector is placed roughly $2\times 10^5 nm$ from the dimer (can be adjusted 
later for testing), with the position of the detector being defined by the polar 
and azimuth angles ($\theta$ \& $\phi$ respectively) such that:
\begin{align}
	[x_{fiber}, y_{fiber}, z_{fiber}] = [rcos(\phi)sin(\theta), rsin(\phi)sin(\theta), rcos(\theta)]
\end{align}
So if you were to place the fibres in the trapping frames x-y plane (i.e. flat against
the cover slip in a covenantal trapping context), in the probe beam frame will only 
have x and z components. The orientation of the detector is just the -ve of its 
position and not scaled by the distance term r, the orientation being the vector from 
the origin of the detector to the center of the dimer.

Pixels can be thought of as points lying on the surface of the detector, by getting 
the spherical coordinates of each pixel we can compute the intensity at each point on 
the detector, allowing us to determine the average intensity on a given detector.

@@

This allows us to define the surface of any detector perpendicular to the origin of 
our coordinate system, we can freely choose the size and resolution of our fibre's. 
For every pixel we define the spherical coordinates relative to our origin, this is 
added to a list of $\theta, \phi$ pairs that are evaluated by \textit{mstm}.

\section{Interpretation of scattering data into orientation estimates}
Consider a dimer in the optical trap (Fig.~\ref{fig:dimer}a), we can define at any 
point in time a unit vector $\hat{s}$ pointing from the centre of the larger sphere 
to the centre of the smaller sphere. A plane wave probe beam is incident on 
the trapping laser, is incident on the dimer, generating a scattering pattern
dependent on the dimer's orientation $I(\hat{\bf s}, \theta, \phi)$ which can be 
computed using \textit{mstm} \cite{I.Mishchenko1996}. To represent the experimental 
set up consisting of a set of optical fibres recording scattered light, we choose 
four sets of spherical angles [$(\theta_1,\ \phi_1), \ (\theta_2,\ \phi_2), \ 
(\theta_3, \ \phi_3), \ (\theta_4,\ \phi_4)$] and record the calculated intensity 
at each angle $I(\hat{s},\ \theta_k,\ \phi_k)$.

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig2a.png}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig2b.png}
	\end{subfigure}
	\caption{(a)Example dimer in orientation $\hat{\bf s}$, (b) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point}
	\label{fig:dimer}
\end{figure}

Our goal is to determine the orientation of the trapped dimer based on the measured 
intensity $I(\hat{n}, \ \theta_k)$. Rather than aim immediately for an exact estimate 
of the dimer's orientation, for the purposes of interpretation of the scattering and 
optimisation of the measurement setup it is more convenient to discretize the possible orientation space into a number of possible reference orientations, which we can then 
use as 'classification categories' in a neural network methodology to map scattering 
data to orientation (see below for further discussion).  Here we choose $\textit{n}_{ref} 
\ = \ 30$ reference orientations $\hat{\bf n}_{\alpha}$  evenly distributed on a 
unit sphere \cite{Reyuthor2006} (Figure~\ref{fig:dimer}b) leading to a maximum 
nearest-neighbour spacing between two neighbouring reference orientations of 0.895 
radians. Using MSTM we compute the raw intensities at each of the measurement angles 
that would be generated by a dimer in each reference orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. While the number and position of detection fibres is technically arbitrary 
there are several constraining factors that limit our ability to infer useful information 
from the trapped object, see Section~\ref{sec:detectors} for a detailed breakdown 
of our choice of detection angles. The raw intensities are normalized according to:
\begin{align}
	\label{eq:scale}
	y_k(\hat{\bf n}_\alpha)
	= 
	\frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
	{\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
\end{align}

where the denominator is simply the standard deviation across the set of values
$I(\hat{\bf n}, \theta_k)$. The reference orientations, raw intensities, and scaled 
signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.

Note that the collected scattering signals are not necessarily simply related to 
their associated reference orientations: as is well known from such examples of 
the inverse scattering problem. While it is trivial to compute the light scattering 
pattern for any given particle with any particular characteristic (i.e. size, 
shape, or orientation), inferring the light scattering from a unknown particle 
to determine said characteristic is incredibly difficult due to complex mapping 
between scattering and said characteristic. Even if the orientation space is divided 
evenly between reference orientation the subsequent signal space ends up being 
appearing mixed making simple comparisons of signals useless for inferring information 
on the particle. Shown below is two clusters of orientation vectors and there 
respective measured scattering signals - the points have been coloured based on 
their proximity to the centre of their respective cluster. While the orientation 
space appears tightly packed and ordered the signal space quickly spreads out in 
an asymmetric fashion. Furthermore as seen in Fig~\ref{fig:mixing}b the signal 
mapping can intersect itself which only further increases the complexity. While 
in some instances the mapping between one reference orientation and another is 
discrete, in other instances the mapping becomes far more complex to discern.

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig3a.png}
	\end{subfigure}
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig3b.png}
	\end{subfigure}
	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals. Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
	\label{fig:mixing}
\end{figure}

Nevertheless, at least where the uncertainty in signal measurements is low (see 
below), we can predict the orientation from the scattering by utilising 
computational techniques such as neural networks. We thus utilised the Python 
machine learning program \textit{scikit-learn} \cite{Pedregosa2011} to build a 
neural network for identifying the dimer's orientation from its light scattering 
signal. The network was trained by generating a database of random orientation 
vectors, calculating the corresponding light scattering signals, and then using 
the network to estimate the probability of a given signal coming from a dimer in 
a given reference orientation. The network's loss function was evaluated and used 
to improve the estimation, the network being trained until the improvement in the 
loss function was less than 0.0001.

Importantly, the estimation provided by the neural network can be improved further 
by accounting for any prior information we know about the dimer, utilising Bayesian 
inference to update the neural network's estimation: 
\begin{align}
	\label{eq:bayes}
	p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}))
	&=
	\frac{p(y_k(\hat{\bf s})|\hat{\bf n}_\alpha)
		p(\hat{\bf n}_\alpha)}{p(y_k(\hat{\bf s}))}
\end{align}

where $p(\hat{\bf n}_\alpha)$ and $p(y_1, y_2, y_3)$ are the prior estimates of 
the distributions of particle orientations and instantaneous signals, respectively.
\textit{Without} any prior evidence we must assume that the orientation prior of 
the dimer $p(\hat{\bf n}_{\alpha})$ is uniform. However, inference about the 
dimer's possible current orientation from knowledge of previous measurements can 
be used to inform our estimate of $p(\hat{\bf n}_{\alpha})$. The latter prior 
$p(y)$ is the probability of measuring a signal ($y_1$, $y_2$, $y_3$). This is 
given by taking the discrete integral over the collection of reference orientations:
\begin{align}
	p(y_1, y_2, y_3, y_4)
	=
	\sum_{\alpha=1}^{n_{\rm ref}}
	p(y_1, y_2, y_3, y_4|\hat{\bf n}_\alpha)
	p(\hat{\bf n}_\alpha)
\end{align}

From \eqref{eq:bayes} we obtain the key result, a mass probability distribution 
denoting the probability that our dimer is in orientation $\hat{\bf n}_{\alpha}$ 
given a measured signal ($y_1$, $y_2$, $y_3$), \textit{i.e.} an estimated mapping 
from scattering measurement to orientation estimate.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculation of error}
\label{sec:divergence}
To evaluate the above estimation of dimer orientation from scattering signal, we use a Brownian simulation of a dimer in the optical trap (Section~\ref{sec:brownian}) to compare estimated most probable reference orientation, derived from the dimer's scattering through Eq.~\eqref{eq:bayes}, with the dimer's known \emph{actual} orientation $\hat{\bf s}$. MSTM provides calculated light scattering from the simulated dimer $I(\hat{\bf s}, \theta)$ and we use \eqref{eq:scale} to obtain normalized values at each measurement angle $\theta_k$,  $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$, from which we obtain $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$. Because we know the actual orientation $\hat{s}$ we can measure the error in the model's estimate by comparing the reference orientation closest to $\hat{s}$, denoted as $\hat{\bf n}_{best}$, with the most probable predicted orientation from Eq.~\eqref{eq:bayes}. An ideal result would be one where the probability distribution is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$:
\begin{align}
	\label{eq:best}
	p_{best} = 
	\begin{cases}
		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
		0 & \text{anywhere else}
	\end{cases}
\end{align}
In reality the distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation, leading to some level 'confidence' in orientation prediction, which can be quantified by calculating the Kullback-Leibler divergence $K_l$ between the two distributions:
\begin{align}
	K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
	= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
	\right]
	\label{eq;kullback}
\end{align}
where a larger value of $K_l$ indicates that our model is less confident in its
prediction of the dimer's orientation. The divergence $K_l$ thus illustrates the 'spread'
in the estimated dimer orientation probability --- a distribution strongly peaked at 
some value would give us more confidence in that value than a near-uniform distribution 
where the scattering measurement could imply a wide range of possible orientations --- 
but it does not directly indicate our estimates actual accuracy, that can be simply defined as the percentage of our estimations that are correct.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Brownian Simulation}
\label{sec:brownian}

We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020Brownian_OT} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap. Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
shaped sphere clusters. We simulate the motion of a dimer trapped in a highly focused Gaussian beam by calculating the optical forces imparted by the laser, and the Brownian force due to the surrounding fluid. MSTM provides the necessary T-matrix to compute the optical force via \textit{ott}. The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}. We simulated a polystyrene dimer ($n = 1.59$) in a suspension of water ($n_{med} = 1.33$) over the course of $1 \ s$ with a simulation time step of $1 \times 10^{-5} \ s$. We placed the dimer 4 microns below the trap focus at an angle $30^{\circ}$ from the horizontal, the resulting trajectory is shown below in Sec~\ref{sec:motion}. We chose these initial parameters because it demonstrates our model's performance in non steady state conditions.

The exact number of detectors was initially assumed to be arbitrary, in that it made no difference to our estimate whether we used 2 angles or 200. For practical purposes it seemed beneficial that we demonstrate our method works for a minimal number of detection angles, as geometric constraints come into play when trying to install a high number of detection fibres for any optical tweezer set up.

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig4a.png}
	\end{subfigure}
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig4b.png}
	\end{subfigure}
	\caption{Expected scattered signals from reference orientations -see fig~\ref{fig:dimer} - when: (a) all three detectors are in the X-Y plane, (b) when 1 detector is raised out of the X-Y plane.}
	\label{fig:detectors}
\end{figure}

When all of the detectors lie in the same plane the expected signal can appear identical despite the dimer being in completely different orientations. This is shown in Figure~\ref{fig:detectors} which plots the expected signals from 30 reference orientations, each point is labelled with its corresponding reference orientation, the fact that points have multiple labels is because the dimer's scattering is indistinguishable in these two reference orientation. It should be noted that these pairs are reflected in one or more axis which suggests that these are due to the arrangement of our detectors. More specifically, if the detectors are placed say in the x-y plane then only when the dimer is pointed nearly fully upright will the expected signal be entirely unique. This is illustrative of the difficulty behind the inverse light scattering problem; as one cannot always map a given signal to a particular parameter value.

To remedy this we raise the third detector out of the x-y plane; as such the expected signals from each reference orientation is unique. As seen between Figures \ref{fig:detectors}a \& b each reference orientation now has a unique scattering signal, though with only three detectors the difference in expected signals can appear insignificant. By adding a $4^{th}$ detector we can differentiate signals more reliably, improving the neural networks performance.   In line with our goal of making this method viable in a laboratory setting we decided not to increase the number of detectors further than 4.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Testing the Model}
\label{sec:test}
Using our simulation from Section~\ref{sec:brownian} we simulated the motion of a silica dimer ($n = 1.45$) trapped in water ($n = 1.33$) within a 5 mW optical trap. The trapping laser is 1064nm NIR focused through a 1.25 NA objective. The dimer is comprised of two tangent spheres with radii $1 \mu m$ and $0.5 \mu m$ respectively. We simulated the first 10 seconds of motion, calculating the orientation and position every 1 ms.

We applied Eq.~\eqref{eq:bayes}, taking the reference orientation with the highest probability  as our estimate of the dimer's instantaneous orientation $\hat{\bf n}_{est}$. To visualise the model's performance  we plotted the radial distance between our estimation $\hat{\bf n}_{est}$ and the dimer's \emph{actual} instantaneous orientation $\hat{\bf s}$ versus time. For comparison, we also plotted the radian distance between the  dimer's instantaneous orientation and the closest reference orientation, denoted $\hat{\bf n}_{best}$. The dotted line indicates the maximum radian distance ($0.896$ radians) between two \textit{neighbouring} reference orientations: if we are under this line then we know our estimate is at least neighbouring the best result. Assuming a uniform prior of the reference orientations $p(\hat{\bf n}_\alpha)$  the neural network's predictions ($\hat{\bf n}_{est}$ from Eq.~\eqref{eq:bayes}) are at times reasonable, but there are significant large and random jumps away from the correct result (Fig.~\ref{fig:uniform}).

\begin{figure}[h!]
	\centering
	\includegraphics[width=\textwidth]{fig5.png}
	\caption{Model's estimation of dimer orientation over the simulation time, assuming uniform prior $p(\hat{\bf n}_\alpha)$, broken up into x, y, and z components for clarity. Blue line denotes the best result we can achieve (the reference orientation $\hat{\bf n}_{best}$ that is closest to the actual orientation), orange line denotes the result provided by eq~\ref{eq:bayes}: where the orange line is not visible, the model's prediction agrees with $\hat{\bf n}_{best}$. Dotted black line is the instantaneous orientation $\hat{\bf s}$.}
	\label{fig:uniform}
\end{figure}

One reason we observe such large jumps in orientation estimated from scattering signals is that there is no simple correlation between the 'distance in scattering space' between scattering signals from two different orientations, and their separation in orientation space: even a large change in orientation can involve a small change in scattering. Combining this fact with use of a uniform prior, indicating essentially no knowledge of how orientation should behave, there is no constraint on how much estimated orientation can change from time-step to time-step. To improve the estimation we can therefore use knowledge of the physical limitations of the object in the trap and its dynamics, imposing a more physically grounded prior, accounting in this case for the fact that the motion of the dimer is limited due to the trap stiffness. Here the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at each time step as a Boltzmann distribution of the physical distance between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$. Put simply, we are reweighing our estimation based on the size of rotation required, with smaller movements being favoured over large movements:

\begin{align}
	p(\hat{\bf n}_\alpha)
	&=\frac{e^{\beta (\hat{\bf n}_\alpha 
			\cdot \hat{\bf n}_{est}(t-\Delta t))}}
	{\sum_{\alpha=1}^{n_{\rm ref}}
		e^{\beta (\hat{\bf n}_\alpha 
			\cdot \hat{\bf n}_{est}(t-\Delta t)}}
	\label{eq:boltz}
\end{align}
Here $\beta$ is a weighting factor describing the dimer's freedom of motion within the trap. As shown in Figure~\ref{fig:biased} implementation of Eq~\eqref{eq:boltz} helps significantly reduce the large random excursions of estimated orientation away from the 'best' result.

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{fig8a.png}
	\caption{\label{fig:biased}
		%
		Estimation of dimer orientation with $p(\hat{\bf n}_\alpha)$ defined
		by Eq~\eqref{eq:boltz}.  Blue line denotes the best result we can
		achieve, orange line denotes the result provided by eq~\ref{eq:bayes}.
		Dotted black line is the instantaneous orientation $\hat{\bf s}$ (see Section~\ref{sec:Bayes}).
		%
	}
\end{figure}

The simulation data from Section~\ref{sec:brownian} was used to evaluate our model's performance --- covered in Section~\ref{sec:divergence}. By summing the divergence of each measurement across the entire simulation we get an evaluation of how well the model performed in estimating the dimer's orientation. To compare the effects of changing certain parameters on the performance of our model we compare our result of $K_{l,total}$ to a worst case scenario and evaluate how much it improves upon this, denoted as $F(K_l)$:
\begin{align}
	K_{l, \ total} &= \sum\limits_{\# =1}^{timesteps} K_{l,\#}
	\\
	K_{l, \ worst} &= \sum\limits_{\#=1}^{timesteps} \ln \left[\frac{1}{1/n_{ref}} \right]
	\\
	F(K_l) &= \frac{K_{l,\ worst}}{K_{l, \ total}}
\end{align}

The worst case scenario is akin to randomly choosing a reference
orientation at each time step. The greater the value of $F(K_l)$, the
better our model's confidence is in characterising the dimer's
motion. Because our model is dependent on several parameters we need
to a sophisticated method for understanding how these parameters
correlate with $F(K_l)$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Asymmetric dimer dynamics}
\label{sec:motion}
A simulation of a asymmetric dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) 
trapped in an off-axis trapping orientation was used as a test case for our 
model, since it was previously shown that dimers can be optically rotated

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{fig7a.png}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig7b.png}
	\end{subfigure}
	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
	\label{fig:motion}
\end{figure}

In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment. However as seen in Figure~\ref{fig:motion} asymmetric dimers demonstrate dynamics that do not immediately achieve steady state. We chose to use asymmetric dimers as our benchmark due to this fact, as its orientational motion is far more complex than a symmetric dimer. In the future we hope to further investigate the motion of asymmetric dimers. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Accounting for sources of error in light scattering measurements}
When it comes to analysing light scattering from any size particle, error analysis becomes a significant factor. Typically this can be accounted for by averaging over long periods of time to get an assessment of the steady state conditions of the target particle. However in our case where we wish to know the instantaneous orientation, we instead have to rely on our understanding of how uncertainty can effect our model's performance. We identified two areas which are likely sources of error in our estimation: firstly, an incorrect modelling of the target particle, and secondly, signal noise arising from experimental factors. We highlight how we address these areas below. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subparagraph{Impact of incorrect dimer sizing}
\label{sec:lam}

One of the main limitations of our model is that we assume that the dimer being modelled in MSTM is accurate to the dimer being trapped in the optical tweezer. Sizing molecules accurately is a significant challenge for single particle analysis so there is bound to be some uncertainty with the measurements. We ran our model 3 times with the neural net being trained on a dimer of size ratio $1:1.95$, $1:2.00$ and $1:2.05$.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.33\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{fig8a.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig8b.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig8c.png}
	\end{subfigure}
	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=9.456]$, (b) 1:2.05 $[F(K_l)=1.324]$, (c) 1:1.95 $[F(K_l)=1.325]$ ($n_{refs} = 30$)}
	\label{fig:size}
\end{figure}

As can be seen from Fig~\ref{fig:size}  even the slightest change in size ratio makes a very significant difference to the performance of our model. This amounts to just over $100 \ nm$ in the dimer's overall size, yet results in our model being correct from over 90 \% of the time to now as low as 30 \%. This highlights the importance of correctly sizing trapped entities before performing any in depth analysis of the scattering pattern, as even the slightest deviation can have a serious impact. We addressed this by increasing the number of available reference orientations from 30 to 126 (following the same procedure as given by \cite{Rey2006} to evenly space out the coordinates) and increasing the weighting factor in Eq~\ref{eq:boltz}. While this didn't have a significant improvement on the overall accuracy of the model, in the worst case having a slight increase from 30.5 \% to 40.3 \%, it did help to significantly reduce the magnitude between our model's estimations and the dimer's motion as seen below in Fig~\ref{fig:refs}.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{fig9a.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig9b.png}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig9c.png}
	\end{subfigure}
	\caption{Model estimates of orientation when neural net has been trained on dimer of size ratio: (a) 1:2 $[F(K_l)=11.756]$, (b) 1:2.05 $[F(K_l)=1.233]$, (c) 1:1.95 $[F(K_l)=2.128]$, ($n_{refs} = 126$)}
	\label{fig:refs}
\end{figure}

Notably the increasing the number of reference orientations had a greater effect when our neural network was trained on a 1:1.95 dimer than a 1:2.05 dimer. This suggests that overshooting our size estimate will be less detrimental to our estimation. Notably if the our sizing is off the neural network does not predict a smooth motion within the trap; instead predicting that the dimer is jumping back and forth between different orientations. This suggest that we can narrow down our estimate of the particle's size by assessing how the dimer is reorienting within the trap, as we should expect a smooth continuos prediction. Since we are working with a spherical dimer it also stands to reason that techniques such as image analysis could be used in part to address this, so long as the trapped entity is sufficiently illuminated. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subparagraph{Impact of measurement noise on model predictions}
\label{sec:epsilon}

So far a key assumption of the neural network implementation is that the detected scattering signal has no uncertainty associated with it. In reality of course scattering signals will always have some non-zero measurement noise. This can be attributed to a variety of factors, from a measurement bias in the detector, to the Brownian motion of the dimer itself. To explore the impact of measurement uncertainty on orientation estimation model performance we introduce a Gaussian noise to the measured signal:
\begin{align}
	I(\hat{\bf s}) = I(\hat{\bf s}) \pm \epsilon I(\hat{\bf s})
\end{align}
where $\epsilon$ is the percentage error associated with the scattering signal. Figure~\ref{fig:epsilon} shows the performance of the model at a range of $\epsilon$ using in-plane detector angles $15^{\circ}$, $55^{\circ}$, $90^{\circ}$ and out-of-plane detector at $75^{\circ}$, with $\beta$ set to $1$:

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig10a.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig10b.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig10c.png}
	\end{subfigure}
	\caption{Model prediction for signal error of (a) $1\%$ ~  $[F(K_l)=7.246]$, (b) $15\%$ $[F(K_l)=0.511]$, and (c) $25\%$ ~  $[F(K_l)=0.536]$.}
	\label{fig:epsilon}
\end{figure}

As can be seen from Figure~\ref{fig:epsilon}, the inclusion of signal noise quickly leads to a decrease in the model's performance. This is due to an inherent feature of the inverse scattering problem: two distinct regions in orientation space can become heavily intertwined and thus no longer well separated when mapped to intensity space (even though the mapping remains continuous): so even small uncertainties in the scattering data can lead to large 'mistakes' in the choice of orientation by the neural network. (Indeed if this was not the case the inverse scattering problem would be quite simple.)

To reduce the effects of the signal noise we took the time average of the expected signal over $0.001 s$ and then had our neural network estimate the orientation based on the average signal.

\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig11a.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig11b.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig11c.png}
	\end{subfigure}
	\caption{Model prediction for signal error of (a) $1\%$ $[F(K_l)=4.823]$,  (b) $15\%$ $[F(K_l)=1.494]$, and (c) $25\%$ $[F(K_l)=0.882]$, time averaged over $1\,{\rm ms}$}
	\label{fig:time average}
\end{figure}
This resulted in a reduction in the overall signal noise and provided a higher degree of accuracy for our model. There appears to be no clear correlation between the length over which we time average and the performance of our model. Time averaging over every 0.05s resulted in a drastically worse performance; this is due to the fact that over longer time periods there is greater uncertainty regarding how the dimer's orientation has changed, thus tracking the instantaneous orientation becomes harder for the neural network. Fortunately, time averaging even over $1\,{\rm ms}$ seems to provide a satisfactory estimation of the dimer's angular dynamics within the optical trap.
